---
title: "Case Study:Cyclistic"
author: "Omegea Hunter"
output: html_notebook
---


Load the libraries necessary to read the CSV's of data from Divvy Trip and Stations and  make columns consistent and merge them into a single dataframe.
```{r Library setup and merge datasets}
# Load necessary packages
install.packages("tidyverse")
install.packages("dplyr")
install.packages("readr")
install.packages("lubridate")
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(purrr)
library(forcats)
library(plyr)

library(scales) # visualization
library(corrplot) # visualization
library(GGally) # visualization
library(ggthemes) # visualization
library(ggalt) # encircle
library(maps) #maps
library(treemap)
library(ggdendro) # Deprogram

# Interactivity
library(crosstalk)
library(plotly)

#Date
library(scales)
library(zoo)
library(lubridate)


`# Set the directory containing the CSV files
directory <- "~/Documents/GitHub/urbanjunglegoddess/projects/data-analytics/How Does a Bike-Share Navigate Speedy Success?/2020-2023 dataset"

# Get the list of CSV files in the directory
file_list <- list.files(directory, pattern = "\\.csv$", full.names = TRUE)

# Initialize an empty dataframe to store the merged data
divvy_trips <- data.frame()

# Loop through each CSV file and merge the data
for (file in file_list) {
  # Read the CSV file
  data <- read_csv(file)
  
  # Merge the data with the existing divvy_trips data frame
  divvy_trips <- rbind(divvy_trips, data)
}
# Import the divvy_stations csv 
divvy_stations_ <- read_csv("Datasets/2017/Divvy_Trips_2017_Q3Q4/Divvy_Stations_2017_Q3Q4.csv", 
    col_types = cols(latitude = col_number(), 
        longitude = col_number(), dpcapacity = col_integer(), 
        online_date = col_datetime(format = "%m/%d/%Y %H:%M"), 
        ...8 = col_skip()))
View(divvy_stations_)

# View the merged data
View(divvy_trips)
colnames(divvy_trips)
colnames(divvy_stations_)
head(divvy_trips)
head(divvy_stations_)
```

Add columns into the data frame named: start_date, start_time, stop_date,
stop_time,ride_length,season(time of the year),and day_of_the_week. 
```{r Adding columns }

# Extract season based on the month of the start_date
 
   divvy_trips$season <- case_when(
    month(divvy_trips$start_date) %in% c(12, 1, 2) ~ "Winter",
    month(divvy_trips$start_date) %in% c(3, 4, 5) ~ "Spring",
    month(divvy_trips$start_date) %in% c(6, 7, 8) ~ "Summer",
    month(divvy_trips$start_date) %in% c(9, 10, 11) ~ "Fall",
    TRUE ~ "Unknown"
    )

# Extract day_of_the_week from start_date
 
    divvy_trips$day_of_the_week <- weekdays(divvy_trips$start_date)

# Calculate the ride_length in minutes
  
    divvy_trips$ride_length <- as.integer(difftime(divvy_trips$stop_time, divvy_trips$start_time, units = "mins"))

# View the modified dataframe
    
    View(divvy_trips)

```

Clean up and add data to prepare for analysis.
```{r Create a new dataframe to clean against the main dataframe}

#Create a new dataframe
divvy_trips_v2 <-divvy_trips

#Remove rows that have information missing 
divvy_trips_v2 <- drop_na(divvy_trips_v2)

#Remove Duplicates 
divvy_trips_v2 <- unique(divvy_trips_v2)
```


Find the outliers
```{r Find the oultiers in the columns}
#Fix the rides that are not a positive number
divvy_trips_v2 <- divvy_trips_v2[divvy_trips_v2$ride_length >= 0,]

```


Changing the data types

```{r changing the data types}
#Change the start_station_id into a integer 
divvy_trips$start_station_id <- as.integer(divvy_trips$start_station_id)

#Change the end_station_id into a integer 
divvy_trips$end_station_id <-as.integer(divvy_trips$end_station_id)

```


  Handle missing values 
    
Find in the start_station_name by their longitude and latitude with the start_lat and the start_lat using the divvy_stations dataset(Use the divvy_trips dataset)
```{r Find start_station_id by lat and log}
    # Filter divvy_trips based on matching start_lat/start_lng in divvy_stations
matched_stations <- divvy_trips %>%
  inner_join(divvy_stations, by = c("start_lat" = "latitude", "start_lng" = "longitude")) %>%
  select(start_lat, start_lng, name, id)

# Preview the matched stations data frame
head(matched_stations)


#See what stations dont have a match 
# Perform a left join between divvy_trips and divvy_stations based on latitude and longitude
matched_stations <- divvy_trips %>%
  left_join(divvy_stations, by = c("start_lat" = "latitude", "start_lng" = "longitude")) %>%
  select(start_lat, start_lng, name = name, id)

# Filter rows where there is no match
unmatched_stations <- matched_stations %>%
  filter(is.na(name) | is.na(id))

# Preview the unmatched stations data frame
head(unmatched_stations)


#Join the divvy_station and divvy_trips data frames by the starting latitutde and longitude of the trip 

divvy_trips_v2 <- divvy_trips %>%
  left_join(divvy_stations, by = c("start_lat" = "latitude", "start_lng" = "longitude"))
# Preview the unmatched stations with match data frame
head(unmatched_with_match)

```
    
Find in the end_station_name by their longitude and latitude with the end_lat and the end_lat using the divvy_stations dataset(Use the divvy_trips dataset)
```{r Find end_station_id by lat and log }

```
    
Correlate station ids the correct stations(Use the divvy_trips dataset)
```{r Correcting station ids}
# Add stations that maybe in divvy_trips and not in the divvy_stations

# Add the correct ids to start_station_id

# Add the correct ids to end_station_id

```
    
Transforming variables
```{r Changing and converting vectors and varibles}
# Convert the start_station_id from characters to integer

# Convert the end_station_id from characters to integer

# Convert started_at from character to time stamp

# Convert ended_at from character to time stamp 


```



Run calculations in one file to get a better sense of the data layout.



Calculate the mean of ride_length 
    ```{r Average of ride length}
    # Calculate the mean of ride_length
        
    mean_ride_length <- mean(divvy_trips_v2$ride_length)

    # Print the mean
      
    print(mean_ride_length)
    
     cat("The average ride length is:", mean_ride_length, "minutes")
    ```
    
Calculate the min ride_length
    ```{r Minimum ride length}
    # Calculate the minimum value of the ride_length column
    
    min_ride_length <- min(divvy_trips_v2$ride_length)

    # Print the minimum ride_length
    
    print(min_ride_length)
    
    cat("The shortest ride is:", min_ride_length, "minutes")
    ```
    
Calculate the max ride_length
    ```{r Maximum ride length}
    # Calculate the maximum value of the ride_length column
    
    max_ride_length <- max(divvy_trips_v2$ride_length)

    # Print the maximum ride_length
    
    print(max_ride_length)
    
    cat("The longest ride is:", max_ride_length, "minutes")
    ```
   
Calculate the mode of day_of_week
    ```{r Most Frequent ride length}
    

    # Print the most frequent ride_length(s)
    
    cat("The most frequent ride_length(s) is/are:", most_frequent_ride_length, "minutes")
    ```
      
     

Conduct descriptive analysis.

Find the basic statistics of the dataframe 

Summaries of the data
```{r The exhaustive summary of divvy_stations_}
#Summary of the whole dataframe
summary(divvy_stations_)

```

```{r The exhaustive summary of divvy_trips}
#Summary of the whole dataframe
summary(divvy_trips)
summary(divvy_trips_v2)
summary(divvy_stations)
summary(divvy_stations_v2)
```

```{r The exhaustive summary of the dataframe}
#Summary of the whole dataframe

#Summary of ride_length

ride_length_summary <- divvy_trips_v2 %>%
  summarise(
    total_rides = n(),
    average_ride_length = mean(ride_length),
    max_ride_length = max(ride_length),
    min_ride_length = min(ride_length),
    median_ride_length = median(ride_length)
  )



busiest_day <- divvy_trips_v2 %>%
  count(started_at) %>%
  filter(n == max(n))

# Display the busiest day
busiest_day



```

```{r The summary of the specific years}
#Summary of the year 2021
#Summary of the year 2022
#Summary of the year 2023
```

```{r The summary of the specific months within the years}
#Summary of all the months in year 2021
#Summary of all the months in year 2022
#Summary of all the months in year 2023
```

```{r The summary of the specific seasons within the specific years}
#Summary of the seasons in the year 2021
#summary of the seasons in the year 2022
#Summary of the seasons in the year 2023
```
  
  Analysis questions
  
What were the annual members doing 
```{r Count of annual members}
#Calculate the amount of annual members

annual_member <- length(which(divvy_trips_v2$member_casual=="member"))

print(annual_member)

```
```{r}

#table(divvy_trips_v2$member_casual)


#nrow(divvy_trips_v2[divvy_trips_v2$start_date > 2020-12-31 & divvy_trips_v2$start_date < 2022-01-01 & divvy_trips_v2$member_casual == "member"])


```


```{r Annual members rides each of the years 2021 to 2023}
#Calculate the amount of annual members in the 2021

#Calculate the amount of annual members in the 2022

#Calculate the amount of annual members in the 2023


#Data viz 
```

```{r Annual members rides each season from 2021 to 2023}
#Calculate the amount of annual members in the specific seasons

#Data viz 
```

```{r Annual members rides each month of the years 2021 to 2023}
#Calculate the amount of annual members in the specific months in sepcific years 

#Data viz 
```

What were the Casual members doing 
```{r Count of Casual members}
#Calculate the amount of Casual members 
casual_member <- length(which(divvy_trips_v2$member_casual=="casual"))

print(casual_member)
```

```{r Casual members rides each of the years 2021 to 2023}
#Calculate the amount of Casual members in the specific years

#Data viz 
```

```{r Casual members rides each season from 2021 to 2023}


#Data viz 
```

```{r Casual members rides each month of the years 2021 to 2023}
#Calculate the amount of Casual members in the specific months in sepcific years 

#Data viz 
```

Who frequented a ride most: member or casual? 
```{r Who showed up the most}
#Calcuate who between the Casual members and the annual memebers rode the most between 2021 to 2023 

 #Data Visualization

```     

What was the most frequented station
```{r The most popular station}

#Data Visualization
```
      
What was the least frequented station
```{r The Least popular station}

#Data Visualization
```



Export a summary file for further analysis.