---
title: "Case Study:Cyclistic" 
author: "Omegea Hunter" 
output: html_document
---
# Case Study: Cyclistic
Cyclistic, a bike-share company in Chicago, wants to understand how casual
riders and annual members use their bikes differently to design a new marketing
strategy. The goal is to convert casual riders into annual members and maximize
the number of memberships for future growth. The project will last for four
weeks and be managed by Lily Moreno, the director of marketing.
The project plan for analyzing the difference between annual members and casual
riders includes creating a Cyclistic database, cleaning data in RStudio,
completing calculations using packages like purrr and forcats, creating
visualizations in Tableau, and presenting findings to stakeholders. The plan
also includes specific deliverables such as an analysis report, recommendations,
and a presentation.
Load the libraries necessary to read the CSV's of data from Divvy Trip and
Stations and make columns consistent and merge them into a single data frame.
```{r Library setup}
#Load necessary packages
install.packages("tidyverse")
install.packages("readr")
install.packages("dplyr")
install.packages("tidyr")
install.packages("ggplot2")
install.packages("tibble")
install.packages("purrr")
install.packages("forcats")
install.packages("plyr")
install.packages("hms")
install.packages("scales")
install.packages("corrplot")
install.packages("GGally")
install.packages("ggthemes")
install.packages("ggalt")
install.packages("maps")
install.packages("leaflet")
install.packages("plotly")
install.packages("lubridate")
install.packages("htmlwidgets")
install.packages("DBI")
install.packages("RMySQL")
install.packages("stringdist")
install.packages("geosphere")
#install.packages("sf")
library(DBI)
library(RMySQL)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(purrr)
library(forcats)
library(plyr)
library(hms)
library(scales)
library(corrplot)
library(GGally)
library(ggthemes)
library(ggalt)
library(maps)
library(leaflet)
library(plotly)
library(scales)
library(lubridate)
library(stringdist)
library(geosphere)
#library(sf)
```
```{r Connect the database}
con <- dbConnect(RMySQL::MySQL(),
                 dbname = "Cyclistic",
                 host = "localhost",
                 port = 3306,
                 user = "root",
                 password = "empirebuilt25!")
divvy_trips <- dbGetQuery(con, "SELECT * FROM divvy_trips")
divvy_stations <- dbGetQuery(con, "SELECT * FROM divvy_stations")
dbDisconnect(con)
```
```{r}
# Set the directory containing the CSV files
directory <- "~/Documents/GitHub/cyclistic_bike_share/2020-2023 dataset"
# Get the list of CSV files in the directory
file_list <- list.files(directory, pattern = "\\.csv$", full.names = TRUE)
# Initialize an empty data frame to store the merged data
divvy_trips <- data.frame()
# Loop through each CSV file and merge the data
for (file in file_list) {
  # Read the CSV file
  data <- read_csv(file
                   , col_types = cols(
    ride_id = col_character(),
    start_station_id = col_integer(),
    end_station_id = col_integer(),
    start_lat=col_double(),
    start_lng=col_double(),
    end_lat=col_double(),
    end_lng=col_double()
    ))
    # Add more columns and their data types as needed
  # Merge the data with the existing divvy_trips data frame
  divvy_trips <- rbind(divvy_trips, data)
}
# Import the divvy_stations csv
divvy_stations<- read_csv("Divvy_Stations_2017_Q3Q4.csv",
    col_types = cols(id = col_integer(),
        dpcapacity = col_integer()))
View(divvy_stations)
rm(data, directory,file,file_list)
# View the merged data
View(divvy_trips)
colnames(divvy_trips)
colnames(divvy_stations)
head(divvy_trips)
head(divvy_stations)
```
## Prepare the data
Add columns into the data frame named: start_date, start_time, stop_date,
stop_time,ride_length,season(time of the year), day_of_the_week
ride_time_of_day.
```{r Convert the date and time columns to proper datetime format}
#Convert the date and time columns to proper datetime format
divvy_trips$started_at <- ymd_hms(divvy_trips$started_at)
divvy_trips$ended_at <- ymd_hms(divvy_trips$ended_at)
# Extract month and year from started_at column
divvy_trips$month <- month(divvy_trips$started_at)
divvy_trips$year <- year(divvy_trips$started_at)
# Remove the online_date column from divvy_stations
divvy_stations <- divvy_stations %>%
  select(-online_date)
#Change member_casual to membership
divvy_trips <- divvy_trips %>% rename(membership = member_casual)
#Calculate the ride_length (duration of the ride) by subtracting the started_at from ended_at
divvy_trips$ride_length <- as.numeric(difftime(divvy_trips$ended_at, divvy_trips$started_at, units = "mins"))

#Find the season the rride was in 
divvy_trips$season <- case_when(
    month(divvy_trips$started_at) %in% c(12, 1, 2) ~ "Winter",
    month(divvy_trips$started_at) %in% c(3, 4, 5) ~ "Spring",
    month(divvy_trips$started_at) %in% c(6, 7, 8) ~ "Summer",
    month(divvy_trips$started_at) %in% c(9, 10, 11) ~ "Fall",
    TRUE ~ "Unknown"
    )
# Extract day_of_the_week from start_date
   divvy_trips$day_of_the_week <- weekdays(divvy_trips$started_at)
# Establish morning, afternoon, and night of the day
  divvy_trips$ride_time_of_day <- case_when(
    hour(divvy_trips$started_at) %in% c(00, 01, 02, 03,04,05,06,07,08,09,10,11) ~ "Morning",
    hour(divvy_trips$started_at) %in% c(12,13,14,15,16) ~ "Afternoon",
    hour(divvy_trips$started_at) %in% c(17,18,19,20,21,22,23) ~ "Night",
    TRUE ~ "Unknown"
    )
 #Create start_location and end_location as spatial points
# divvy_trips$start_location <- st_point(cbind(divvy_trips$start_lng, divvy_trips$start_lat))
# divvy_trips$end_location <- st_point(cbind(divvy_trips$end_lng, divvy_trips$end_lat))
  
# View the modified data frame
    View(divvy_trips)
```
## Clean Data
```{r Create a new dataframe to clean against the main dataframe}
#Create a back up data frames
divvy_trips_v2 <-divvy_trips
divvy_stations_v2 <- divvy_stations
#Remove Duplicates
divvy_trips_v2 <- unique(divvy_trips_v2)
#Calculate the total number of trips in the divvy_trips
total_trips <- nrow(divvy_trips)
# Count of each rideable type in the divvy_trips
rideable_counts <- table(divvy_trips$rideable_type)
# Extract records with missing data from divvy_trips_v2
records_missing_info <- divvy_trips_v2 %>%
  filter_all(any_vars(is.na(.)))
# Extract unique start station information
start_stations <- divvy_trips %>%
  distinct(start_station_name, start_station_id, start_lat, start_lng) %>%
  rename(station_name = start_station_name, station_id = start_station_id, latitude = start_lat, longitude = start_lng)
# Extract unique end station information
end_stations <- divvy_trips %>%
  distinct(end_station_name, end_station_id, end_lat, end_lng) %>%
  rename(station_name = end_station_name, station_id = end_station_id, latitude = end_lat, longitude = end_lng)
# Combine start and end station information
all_stations <- bind_rows(start_stations, end_stations)
# Print the resulting data frame
print(all_stations)
# Compare all_stations to divvy_stations by station_id
comparison <- all_stations %>%
  anti_join(divvy_stations, by = "id")
# Print the unmatched station information
print(comparison)
# Compare all_stations to divvy_stations by station_id
matching_stations <- all_stations %>%
  semi_join(divvy_stations, by = "id")
# Print the matching station information
print(matching_stations)
```
Find the outliers
```{r Find the oultiers in the columns}
# Count the number of null values in a column
membership_null_count <- sum(is.na(divvy_trips$membership))
# Find ride lengths less than 0 in divvy_trips_v2
negative_ride_lengths <- divvy_trips_v2 %>%
  filter(ride_length < 0)
#Fix the rides that are not a positive number
divvy_trips_v2 <- divvy_trips_v2[divvy_trips_v2$ride_length >= 0,]
# Check and correct the spelling of membership values
divvy_trips_v2 <- divvy_trips_v2 %>%
  mutate(membership = case_when(
    membership == "member" | membership == "casual" ~ membership,
    membership == "memmber" ~ "member",
    membership == "casua" ~ "casual",
    TRUE ~ NA_character_
  ))
# Find inconsistent data in divvy_trips
inconsistent_trips <- divvy_trips[!(divvy_trips$rideable_type %in% c("classic_bike", "docked_bike", "electric_bike")), ]
# Create a new data frame for inconsistent data
inconsistent_data <- data.frame(inconsistent_trips, inconsistent_stations)
#Remove rows that have information missing
records_missing_info <-divvy_trips_v2 <- drop_na(divvy_trips_v2)
# Find similar words in start_station_name and end_station_name columns of divvy_trips
similar_words_trips <- stringdist::amatch(unique(c(divvy_trips$start_station_name, divvy_trips$end_station_name)), unique(c(divvy_stations$name)), maxDist = 2)
# Find similar words in name column of divvy_stations
similar_words_stations <- stringdist::amatch(unique(divvy_stations$name), unique(c(divvy_trips$start_station_name, divvy_trips$end_station_name)), maxDist = 2)
```
```{r Replacing the empty spaces as much as possible}
# Match the stations
# # Find station names in divvy_trips_v2 that are not in divvy_stations_v2
# missing_start_station_names <- divvy_trips_v2 %>%
#   anti_join(divvy_stations_v2, by = c("start_station_name" = "name")) %>%
#   select(start_station_name, start_station_id, start_lat, start_lng)
#
# missing_end_station_names <- divvy_trips_v2 %>%
#   anti_join(divvy_stations_v2, by = c("end_station_name" = "name")) %>%
#   select(end_station_name, end_station_id, end_lat, end_lng)
#
# missing_start_station_names <- missing_start_station_names %>%
#   rename(id = start_station_id, latitude = start_lat, longitude = start_lng)
#
# missing_end_station_names <- missing_end_station_names %>%
#   rename(id = end_station_id, latitude = end_lat, longitude = end_lng)
#
# # Add missing stations to divvy_stations_v2 data frame
# missing_station_names <- merge(missing_start_station_names, missing_end_station_names, by = "id")
#
# # Remove duplicate stations based on station ID
# missing_station_names <- unique(missing_station_names)
#
# # Move missing station information to divvy_stations_v2 data frame
# divvy_stations_v2 <- rbind(divvy_stations_v2, missing_station_names)
#
# # Print the updated divvy_stations_v2 data frame
# print(divvy_stations_v2)
```


```{r}
cleaned_divvy_trips <- divvy_stations_v2 %>%
  select(
    ride_id,
    rideable_type,
    started_at,
    ended_at,
    start_station_name,
    start_station_id,
    end_station_name,
    end_station_id,
    start_lat,
    start_lng,
    end_lat,
    end_lng,
    membership,
    season,
    day_of_the_week,
    ride_length,
    ride_time_of_day
  )
cleaned_divvy_stations <- divvy_stations_v2 %>%
  select(
    id,
    name,
    city,
    latitude,
    longitude,
    dpcapacity,
  )
```
Data Frame Summaries
```{r}
#Summaries of all the data frames
summary(divvy_trips)
summary(divvy_trips_v2)
summary(divvy_stations)
summary(divvy_stations_v2)
summary(cleaned_divvy_trips)
summary(cleaned_divvy_stations)
#Summary of ride_length
ride_length_summary <- cleaned_divvy_trips %>%
  summarise(
    total_rides = n(),
    average_ride_length = mean(ride_length),
    max_ride_length = max(ride_length),
    min_ride_length = min(ride_length),
    median_ride_length = median(ride_length)
  )
#Calculate the amount of annual members all together
total_annual_members <- cleaned_divvy_trips %>%
  filter(membership == "Annual") %>%
  summarise(total_members = n())
#Calculate the amount of annual members in the 2020
annual_members_2020 <- cleaned_divvy_trips %>%
  filter(membership == "Annual" & year(started_at) == 2020) %>%
  summarise(total_members = n())
#Calculate the amount of annual members in the 2021
annual_members_2021 <- cleaned_divvy_trips %>%
  filter(membership == "Annual" & year(started_at) == 2021) %>%
  summarise(total_members = n())
#Calculate the amount of annual members in the 2022
annual_members_2022 <- cleaned_divvy_trips %>%
  filter(membership == "Annual" & year(started_at) == 2022) %>%
  summarise(total_members = n())
#Calculate the amount of annual members in the 2023
annual_members_2023 <- cleaned_divvy_trips %>%
  filter(membership == "Annual" & year(started_at) == 2023) %>%
  summarise(total_members = n())
#Calculate the amount of annual members in the specific seasons
annual_members_by_season <- cleaned_divvy_trips %>%
  filter(membership == "Annual") %>%
  group_by(season) %>%
  summarise(total_members = n())
#Calculate the amount of annual members in the specific months in specific years
annual_members_by_month <- cleaned_divvy_trips %>%
  filter(membership == "Annual" & year(started_at) %in% c(2020,2021, 2022, 2023)) %>%
  group_by(year(started_at), month(started_at)) %>%
  summarise(total_members = n())
#Calculate the amount of Casual members in the specific years
casual_members_by_year <- cleaned_divvy_trips %>%
  filter(membership == "Casual" & year(started_at) %in% c(2020,2021, 2022, 2023)) %>%
  group_by(year(started_at)) %>%
  summarise(total_members = n())
#Calculate the amount of Casual members in the specific seasons
casual_members_by_season <- cleaned_divvy_trips %>%
  filter(membership == "Casual") %>%
  group_by(season) %>%
  summarise(total_members = n())
#Calculate the amount of Casual members in the specific months in specific years
casual_members_by_month <- cleaned_divvy_trips %>%
  filter(membership == "Casual" & year(started_at) %in% c(2020, 2021, 2022, 2023)) %>%
  group_by(year(started_at), month(started_at)) %>%
  summarise(total_members = n())
```
## Analyze Data
```{r}
# Count the number of rides by member type
ride_counts <- cleaned_divvy_trips %>%
  group_by(membership) %>%
  summarise(ride_count = n())
# Average ride length by member type
ride_avg_length <- cleaned_divvy_trips %>%
  group_by(membership) %>%
  summarise(avg_ride_length = mean(ride_length))
#Calculate the percentage of rides taken by each user type (member or casual) within the years 2020-2023
ride_percentages <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at), membership) %>%
  summarise(ride_percentage = n() / nrow(cleaned_divvy_trips) * 100)
# Calculate correlation matrix
corr_matrix <- cor(cleaned_divvy_trips[, c("ride_length", "start_lat", "start_lng", "end_lat", "end_lng")])
# Plot the number of rides over time
rides_over_time <- cleaned_divvy_trips %>%
  group_by(year, month) %>%
  summarise(total_rides = n())
# Bar chart of ride counts by member type and season
ride_counts_season <- cleaned_divvy_trips %>%
  group_by(membership, season) %>%
  summarise(ride_count = n())
# Top 10 most popular start stations
top_start_stations <- cleaned_divvy_trips %>%
  group_by(start_station_name) %>%
  summarise(ride_count = n()) %>%
  top_n(10, ride_count) %>%
  arrange(desc(ride_count))
# Top 10 most popular end stations
top_end_stations <- cleaned_divvy_trips %>%
  group_by(end_station_name) %>%
  summarise(ride_count = n()) %>%
  top_n(10, ride_count) %>%
  arrange(desc(ride_count))
# Proportion of member types
member_type_counts <- cleaned_divvy_trips %>%
  count(membership)
# Proportion of member types by season
member_type_season_counts <- cleaned_divvy_trips %>%
  count(membership, season)
# Count the number of rides by day of the week
ride_counts_by_day <- cleaned_divvy_trips %>%
  mutate(day_of_week = weekdays(started_at)) %>%
  count(day_of_week)
# Count the number of rides by season
ride_counts_by_season <- cleaned_divvy_trips %>%
  count(season)
# Calculate average ride duration by member type and season
ride_duration_by_member_season <- cleaned_divvy_trips %>%
  group_by(membership, season) %>%
  summarise(avg_ride_duration = mean(ride_length))
#Calculate the average ride length (duration) for each season within the years 2020-2023
average_ride_length <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  mutate(ride_length = ended_at - started_at) %>%
  group_by(year(started_at), season) %>%
  summarise(average_length = mean(ride_length, na.rm = TRUE))
#Calculate the average ride duration for each month within the years 2020-2023:
average_ride_duration <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  mutate(ride_duration = ended_at - started_at) %>%
  group_by(year(started_at), month(started_at)) %>%
  summarise(average_duration = mean(ride_duration, na.rm = TRUE))
#Calculate the total number of rides for each day of the week within the years 2020-2023
total_rides_per_day <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at), day_of_the_week) %>%
  summarise(total_rides = n())
# Calculate ride distances using start and end coordinates
cleaned_divvy_trips$ride_distance <- geosphere::distHaversine(
  p1 = cbind(cleaned_divvy_trips$start_lng, cleaned_divvy_trips$start_lat),
  p2 = cbind(cleaned_divvy_trips$end_lng, cleaned_divvy_trips$end_lat)
)
# Find the busiest day
busiest_day <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(started_at_date) %>%
  summarise(total_rides = n()) %>%
  filter(total_rides == max(total_rides))
# Which location gets busiest on what season in what year
busiest_location_by_season <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at), season, start_station_name) %>%
  summarise(total_rides = n()) %>%
  filter(total_rides == max(total_rides))
# Which is least busiest in every year
least_busy_locations <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at), start_station_name) %>%
  summarise(total_rides = n()) %>%
  filter(total_rides == min(total_rides))
# What season is the busiest
busiest_season <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(season) %>%
  summarise(total_rides = n()) %>%
  filter(total_rides == max(total_rides))
# What year was the busiest
busiest_year <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at)) %>%
  summarise(total_rides = n()) %>%
  filter(total_rides == max(total_rides))
# What was the busiest month in what year
busiest_month_by_year <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at), month(started_at)) %>%
  summarise(total_rides = n()) %>%
  group_by(year(started_at)) %>%
  filter(total_rides == max(total_rides))
# Summary of the busiest day
busiest_day_summary <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(started_at_date) %>%
  summarise(total_rides = n())
# Summary of the type of bike used by year
bike_type_summary <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(year(started_at), rideable_type) %>%
  summarise(total_rides = n())
# Summary of the year 2020
summary_2020 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2020-12-31")) %>%
  summarise(total_rides = n())
# Summary of the year 2021
summary_2021 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2021-01-01") & started_at <= as.Date("2021-12-31")) %>%
  summarise(total_rides = n())
# Summary of the year 2022
summary_2022 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2022-01-01") & started_at <= as.Date("2022-12-31")) %>%
  summarise(total_rides = n())
# Summary of the year 2023
summary_2023 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2023-01-01") & started_at <= as.Date("2023-12-31")) %>%
  summarise(total_rides = n())
# Summary of all the months in year 2021
summary_months_2021 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2021-01-01") & started_at <= as.Date("2021-12-31")) %>%
  group_by(month(started_at)) %>%
  summarise(total_rides = n())
# Summary of all the months in year 2022
summary_months_2022 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2022-01-01") & started_at <= as.Date("2022-12-31")) %>%
  group_by(month(started_at)) %>%
  summarise(total_rides = n())
# Summary of all the months in year 2023
summary_months_2023 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2023-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(month(started_at)) %>%
  summarise(total_rides = n())
# Summary of the seasons in the year 2020
summary_seasons_2020 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2020-12-31")) %>%
  group_by(season) %>%
  summarise(total_rides = n())
# Summary of the seasons in the year 2021
summary_seasons_2021 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2021-01-01") & started_at <= as.Date("2021-12-31")) %>%
  group_by(season) %>%
  summarise(total_rides = n())
# Summary of the seasons in the year 2022
summary_seasons_2022 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2022-01-01") & started_at <= as.Date("2022-12-31")) %>%
  group_by(season) %>%
  summarise(total_rides = n())
# Summary of the seasons in the year 2023
summary_seasons_2023 <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2023-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(season) %>%
  summarise(total_rides = n())
#Calculate who between the Casual members and the annual members rode the most between 2020 to 2023
casual_vs_annual <- cleaned_divvy_trips %>%
  filter(started_at >= as.Date("2020-01-01") & started_at <= as.Date("2023-12-31")) %>%
  group_by(membership) %>%
  summarise(total_rides = n())
```
## Visualizations
```{r}
# Histogram of ride lengths
ggplot(cleaned_divvy_trips, aes(x = ride_length)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  labs(x = "Ride Length (minutes)", y = "Count", title = "Distribution of Ride Lengths")
# Bar chart of ride counts by member type
ggplot(ride_counts, aes(x = membership, y = ride_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Member Type", y = "Ride Count", title = "Ride Counts by Member Type")
# Plot correlation matrix
corrplot(corr_matrix, method = "color", type = "upper", tl.col = "black")
# Plot the number of rides over time
ggplot(rides_over_time, aes(x = month, y = total_rides, group = year, color = factor(year))) +
  geom_line() +
  labs(x = "Month", y = "Number of Rides", title = "Number of Rides Over Time") +
  scale_color_discrete(name = "Year")
# Create a leaflet map of station locations
leaflet(cleaned_divvy_stations) %>%
  addTiles() %>%
  addMarkers(lng = ~longitude, lat = ~latitude, popup = ~name)
# Bar chart of ride counts by member type and season
ggplot(ride_counts_season, aes(x = membership, y = ride_count, fill = season)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Member Type", y = "Ride Count", title = "Ride Counts by Member Type and Season") +
  scale_fill_manual(values = c("spring" = "blue", "summer" = "green", "fall" = "orange", "winter" = "purple"))
# Boxplot of ride duration by member type
ggplot(cleaned_divvy_trips, aes(x = membership, y = ride_length)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "Member Type", y = "Ride Duration (minutes)", title = "Ride Duration by Member Type")
# Top 10 most popular start stations
ggplot(top_start_stations, aes(x = reorder(start_station_name, -ride_count), y = ride_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Start Station", y = "Ride Count", title = "Top 10 Most Popular Start Stations") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
# Top 10 most popular end stations
ggplot(top_end_stations, aes(x = reorder(end_station_name, -ride_count), y = ride_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "End Station", y = "Ride Count", title = "Top 10 Most Popular End Stations") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
# Proportion of member types
ggplot(member_type_counts, aes(x = "", y = n, fill = membership)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(x = NULL, y = NULL, fill = "Member Type", title = "Proportion of Member Types") +
  theme_void() +
  theme(legend.position = "bottom")
# Proportion of member types by season
ggplot(member_type_season_counts, aes(x = season, y = n, fill = membership)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(x = "Season", y = "Proportion", fill = "Member Type", title = "Proportion of Member Types by Season")
# Plot the ride frequency by day of the week
ggplot(ride_counts_by_day, aes(x = day_of_week, y = n, fill = day_of_week)) +
  geom_bar(stat = "identity") +
  labs(x = "Day of the Week", y = "Ride Count", title = "Ride Frequency by Day of the Week") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_discrete(name = "Day of the Week")
# Plot the distribution of ride lengths
ggplot(cleaned_divvy_trips, aes(x = ride_length)) +
  geom_histogram(binwidth = 5, fill = "steelblue") +
  labs(x = "Ride Length (minutes)", y = "Count", title = "Distribution of Ride Lengths")
# Plot the ride counts by season
ggplot(ride_counts_by_season, aes(x = season, y = n, fill = season)) +
  geom_bar(stat = "identity") +
  labs(x = "Season", y = "Ride Count", title = "Ride Counts by Season") +
  scale_fill_discrete(name = "Season")
# Plot the average ride duration by member type and season
ggplot(ride_duration_by_member_season, aes(x = membership, y = avg_ride_duration, fill = season)) +
  geom_col(position = "dodge") +
  labs(x = "Member Type", y = "Average Ride Duration", title = "Average Ride Duration by Member Type and Season") +
  scale_fill_discrete(name = "Season")
# Plot the distribution of ride distances
ggplot(cleaned_divvy_trips, aes(x = ride_distance)) +
  geom_histogram(binwidth = 1000, fill = "steelblue") +
  labs(x = "Ride Distance (meters)", y = "Count", title = "Distribution of Ride Distances")
# Bar plot the distribution of rideable types
ggplot(cleaned_divvy_trips, aes(x = rideable_type)) +
  geom_bar(fill = "blue") +
  labs(title = "Distribution of Rideable Types", x = "Rideable Type", y = "Count")
```
Print the Data frame
```{r}
print(divvy_trips)
print(divvy_trips_v2)
print(divvy_stations)
print(divvy_stations_v2)
print(cleaned_divvy_stations)
print(cleaned_divvy_trips)
print(negative_ride_lengths)
print(records_missing_info)
print(total_trips)
print(rideable_counts)
print()
```


